Strict Pure Functional Programming and Automatic Parallelization and Transparent Fault Tolerance\\ (New project paper)

### 2016-03-31 pubs/tfp16/paper.tex
do emacs save first

\subsubsection{\emph{Strict but still pure} is not new.}
Very simple implementations have been prototyped, for example
Sheard's.~\cite{Sheard:2003}
%
A plug-in for GHC (prior to 8.0) exists for making functions, but not data
structures, strict.~\cite{strict-ghc-plugin}.  The transformation is on GHC
Core, recursively expand non-recursive let bindings into case bindings.  Thus
it seems that while it makes functions denotationally strict, function arguments
are still evaluated after function call.
%
The Disciplined Disciple Compiler project appears to intend implementation of
a strict, higher order functional language (or family of languages) compiler
that has been underway for some time, but the emphasis appears to be on a much
more complex type system for handling stateful computation, and the prospects
for its eventual completion are unknown~\cite{disciple}.
%
NVIDIA's NOVA language, and corresponding compiler, implements a strict
(``call by value''), pure, polymorphic, higher-order functional language with
built-in parallel operators (e.g., \emph{map}, \emph{reduce}, and
\emph{scan}).~\cite{NVIDIA:2013} Three back-ends were implemented: for
sequential C, parallel (multi-threaded) C, and CUDA C, the lattermost
targeting NVIDIA GPUs.

>>>>

\subsubsection{\emph{Strict but still pure} is not new.}
Very simple implementations have been prototyped, for example
Sheard's.~\cite{Sheard:2003}
%
A plug-in for GHC (prior to 8.0) exists for making functions, but not data
structures, strict.~\cite{strict-ghc-plugin}.  The transformation is on GHC
Core, recursively expand non-recursive let bindings into case bindings.  Thus
it seems that while it makes functions denotationally strict, function arguments
are still evaluated after function call.
%
The Disciplined Disciple Compiler project appears to intend implementation of
a strict, higher order functional language (or family of languages) compiler
that has been underway for some time, but the emphasis appears to be on a much
more complex type system for handling stateful computation, and the prospects
for its eventual completion are unknown~\cite{disciplined-disciple}.
%
NVIDIA's NOVA language, and corresponding compiler, implements a strict
(``call by value''), pure, polymorphic, higher-order functional language with
built-in parallel operators (e.g., \emph{map}, \emph{reduce}, and
\emph{scan}).~\cite{NVIDIA:2013} Three back-ends were implemented: for
sequential C, parallel (multi-threaded) C, and CUDA C, the lattermost
targeting NVIDIA GPUs.
